Comparison and Evaluation of Popular Syntax Highlighting Themes in Terms of Color Correlated Temperature Values

### 2022-04-23 doc/manuscript.tex
.

\title{Comparison and Evaluation of Popular Syntax Highlighting Themes in Terms of Color Correlated Temperature values}

@@

@@

Based on VS Code theme statistics\cite{vs2022thememarket}, the most popular 17 themes have been chosen. Some of these
themes have dark and light versions.Since the dark theme is mostly used\cite{sarath2016prefer}, the number of examples
of this theme was chosen more than the light theme.Since there is no API to retrieve VS code Market information
programmatically, theme information was obtained for each theme by web scraping. Information on themes selected within
the scope of the study are shown in Table \ref{tab:themes}.

@@

\vspace{0.1in}
\begin{verbbox}
#include "cache.h"
\end{verbbox}
\noindent\hspace{1in}\fbox{\theverbbox}
\vspace{0.1in}

@@

\begin{figure}
  \centering
  \begin{BVerbatim}
    Version:0.9
    StartHTML:0000000105
    EndHTML:0000186320
    StartFragment:0000000141
    EndFragment:0000186284
    <html>
    <body>
    <!--StartFragment--><div><span style="color: #c74ded;">
    #include</span><span style="color: #d5ced9;"> </span>
    <span style="color: #96e072;">"cache.h"</span></div>
    ...
    </body>
    </html>
  \end{BVerbatim}
  \caption{The raw HTML output of colorized sample code}
  \label{code:syntaxhtml}
\end{figure}

@@

\begin{itemize}
\item There is some meta-information at the beginning of the file. These lines must be removed for the content to be a valid HTML file.
\item Each line has been converted into a \texttt{div} blockquote in HTML.
\item Colorings are transformed into color-coded \texttt{span} parts.
\end{itemize}

@@

\begin{figure}[htbp]
  \caption{BDF definition of the Consolas font for the character \texttt{A}}
  \label{code:letter_A_BDF}

  \centering
  \begin{BVerbatim}
    STARTCHAR 0041
    ENCODING 65
    SWIDTH 514 0
    DWIDTH 10 0
    BBX 10 12 0 0
    BITMAP
    0E00
    0E00
    1A00
    1300
    1300
    3100
    3180
    2180
    7F80
    60C0
    40C0
    C0C0
    ENDCHAR
  \end{BVerbatim}
\end{figure}

\begin{figure}[htbp]
  \caption{Bit pattern in Consolas font of character \texttt{A}}
  \label{fig:letter_A_bitmap}

  \centering
  \begin{BVerbatim}
    0 0 0 0 1 1 1 0 0 0 0 0 0 0 0 0 
    0 0 0 0 1 1 1 0 0 0 0 0 0 0 0 0 
    0 0 0 1 1 0 1 0 0 0 0 0 0 0 0 0 
    0 0 0 1 0 0 1 1 0 0 0 0 0 0 0 0 
    0 0 0 1 0 0 1 1 0 0 0 0 0 0 0 0 
    0 0 1 1 0 0 0 1 0 0 0 0 0 0 0 0 
    0 0 1 1 0 0 0 1 1 0 0 0 0 0 0 0 
    0 0 1 0 0 0 0 1 1 0 0 0 0 0 0 0 
    0 1 1 1 1 1 1 1 1 0 0 0 0 0 0 0 
    0 1 1 0 0 0 0 0 1 1 0 0 0 0 0 0 
    0 1 0 0 0 0 0 0 1 1 0 0 0 0 0 0 
    1 1 0 0 0 0 0 0 1 1 0 0 0 0 0 0 
  \end{BVerbatim}
\end{figure}

Considering the bit pattern at Figure \ref{fig:letter_A_bitmap} "Character bit density" is calculated  as follows.

@@

\[
  \textrm{Character bit density} = \frac{44}{{12}\cdot{16}} = 0.229
\]

\def\pixelmapnobb{{%
  {1,1,0,0,0,0,0,0,1,1,0,0,0,0,0,0},
  {0,1,0,0,0,0,0,0,1,1,0,0,0,0,0,0},
  {0,1,1,0,0,0,0,0,1,1,0,0,0,0,0,0},
  {0,1,1,1,1,1,1,1,1,0,0,0,0,0,0,0},
  {0,0,1,0,0,0,0,1,1,0,0,0,0,0,0,0},
  {0,0,1,1,0,0,0,1,1,0,0,0,0,0,0,0},
  {0,0,1,1,0,0,0,1,0,0,0,0,0,0,0,0},
  {0,0,0,1,0,0,1,1,0,0,0,0,0,0,0,0},
  {0,0,0,1,0,0,1,1,0,0,0,0,0,0,0,0},
  {0,0,0,1,1,0,1,0,0,0,0,0,0,0,0,0},
  {0,0,0,0,1,1,1,0,0,0,0,0,0,0,0,0},
  {0,0,0,0,1,1,1,0,0,0,0,0,0,0,0,0},
}}
\def\pixelmapbb{{%
  {1,1,0,0,0,0,0,0,1,1},
  {0,1,0,0,0,0,0,0,1,1},
  {0,1,1,0,0,0,0,0,1,1},
  {0,1,1,1,1,1,1,1,1,0},
  {0,0,1,0,0,0,0,1,1,0},
  {0,0,1,1,0,0,0,1,1,0},
  {0,0,1,1,0,0,0,1,0,0},
  {0,0,0,1,0,0,1,1,0,0},
  {0,0,0,1,0,0,1,1,0,0},
  {0,0,0,1,1,0,1,0,0,0},
  {0,0,0,0,1,1,1,0,0,0},
  {0,0,0,0,1,1,1,0,0,0},
}}

@@

\begin{figure}[htbp]
  \begin{minipage}{0.45\textwidth}
    \centering
    \caption{Consolas TTF '\texttt{A}' letter}
    \label{fig:consolas_A}

    \includesvg[inkscapelatex=false, width=140pt]{consolas_A.svg}
  \end{minipage}\hfill%
  \begin{minipage}{0.48\textwidth}
    \centering
    \caption{Consolas BDF '\texttt{A}' letter}
    \label{fig:letter_A}

@@

    \foreach \x in {0, 1, ..., 15}{
      \foreach \y in {0, 1, ...,11}{
        \def\bit{0} \pgfmathsetmacro{\bit}{\pixelmapnobb[mod(\y,12)][mod(\x,16)]}
        \node[box, fill=color\bit] at (\x,\y){};
      }
    }
    \end{tikzpicture}
  \end{minipage}
\end{figure}

>>>>

\title{Comparison and Evaluation of Popular Syntax Highlighting Themes in Terms of Color Correlated Temperature Values}

@@

\begin{figure}[htb]
  \centering
  \begin{BVerbatim}
    #include "cache.h"
    #include "config.h"
    ···
    static void std_output(struct grep_opt *opt, const void *buf, size_t size)
    {
            fwrite(buf, size, 1, stdout);
    }
  \end{BVerbatim}
  \caption{Sample code without coloring}
  \label{code:string}
\end{figure}

\begin{figure}[h]
  \centering
  \includesvg[inkscapelatex=false, width=400pt]{code.svg}
  \caption{Colored code}
  \label{fig:code}
\end{figure}

\begin{figure}[H]
  \centering
  \includesvg[inkscapelatex=false, width=200pt]{string.svg}
  \caption{Colored code content as a single string}
  \label{fig:string}
\end{figure}

@@

Based on VS Code theme statistics\cite{vs2022thememarket}, the most popular 17 themes have been chosen. Some of these
themes have dark and light versions. Since the dark theme is mostly used\cite{sarath2016prefer}, the number of examples
of this theme was chosen more than the light theme. Since there is no API to retrieve VS code Market information
programmatically, theme information was obtained for each theme by web scraping. Information on themes selected within
the scope of the study are shown in Table \ref{tab:themes}.

@@

\vspace{0.1in}
\begin{BVerbatim}
  #include "cache.h"
  ...
\end{BVerbatim}
\vspace{0.1in}

@@

\begin{figure}[H]
  \centering
  \begin{BVerbatim}
    Version:0.9
    StartHTML:0000000105
    EndHTML:0000186320
    StartFragment:0000000141
    EndFragment:0000186284
    <html>
    <body>
      <!--StartFragment-->
      <div style="color: #d4d4d4;background-color: #1e1e1e;
      font-family: Consolas, 'Courier New', monospace;font-weight: normal;
      font-size: 14px;line-height: 19px;white-space: pre;">
      <div><span style="color: #c586c0;">#include</span>
      <span style="color: #569cd6;"> </span>
      <span style="color: #ce9178;">"cache.h"</span></div>
      ...
    </body>
    </html>
  \end{BVerbatim}
  \caption{The raw HTML output of colorized sample code}
  \label{code:syntaxhtml}
\end{figure}

@@

\begin{itemize}
\item There is some meta-information at the beginning of the file. These lines must be removed for the content to be a valid HTML file.
\item Each line has been converted into a \texttt{div} blockquote in HTML.
\item Colorings are transformed into color-coded \texttt{span} parts.
\item Font family (`Consolas`) and size (`14px`) are defined at the beginning.
\item Line height (`19px`), which will be used to compute a full height character box, is defined.
\item Background color, which will be used for space characters, is defined.
\end{itemize}

@@

\begin{figure}[htbp]
  \caption{Selected fields from BDF font properties section of the Consolas font}
  \label{code:consolas_properties_BDF}

  \centering
  \begin{BVerbatim}
    SIZE 14 100 100
    FONTBOUNDINGBOX 19 27 -6 -6
    FAMILY_NAME "Consolas"
    WEIGHT_NAME "Medium"
    SLANT "R"
  \end{BVerbatim}
\end{figure}

\begin{figure}[htbp]
  \caption{BDF definition of the Consolas font for the character '\texttt{a}'}
  \label{code:letter_A_BDF}

  \centering
  \begin{BVerbatim}
    STARTCHAR 0061
    ENCODING 97
    SWIDTH 514 0
    DWIDTH 10 0
    BBX 8 9 1 0
    BITMAP
    3E
    43
    03
    03
    3F
    43
    C3
    C7
    7B
    ENDCHAR
  \end{BVerbatim}
\end{figure}

\begin{figure}[htbp]
  \caption{Bit pattern in Consolas font of character '\texttt{a}'}
  \label{fig:letter_A_bitmap}

  \centering
  \begin{BVerbatim}
    0 0 1 1 1 1 1 0
    0 1 0 0 0 0 1 1
    0 0 0 0 0 0 1 1
    0 0 0 0 0 0 1 1
    0 0 1 1 1 1 1 1
    0 1 0 0 0 0 1 1
    1 1 0 0 0 0 1 1
    1 1 0 0 0 1 1 1
    0 1 1 1 1 0 1 1
  \end{BVerbatim}
\end{figure}

\begin{comment}
Considering the bit pattern at Figure \ref{fig:letter_A_bitmap} "Character bit density" is calculated  as follows.

@@

\[
  \textrm{Character bit density} = \frac{44}{{12}\cdot{16}} = 0.229
\]
\end{comment}

\begin{figure}[htbp]
    \centering
    \includesvg[inkscapelatex=false, width=250pt]{charbox.svg}
    \caption{Consolas TTF of '\texttt{a}' character with line height}
    \label{fig:charbox}
\end{figure}

\def\pixelmap_a{{%
  {0,0,0,0,0,0,0,0,0,0,0,0,0,0},
  {0,0,0,0,0,0,0,0,0,0,0,0,0,0},
  {0,1,1,1,1,0,1,1,0,0,0,0,0,0},
  {1,1,0,0,0,1,1,1,0,0,0,0,0,0},
  {1,1,0,0,0,0,1,1,0,0,0,0,0,0},
  {0,1,0,0,0,0,1,1,0,0,0,0,0,0},
  {0,0,1,1,1,1,1,1,0,0,0,0,0,0},
  {0,0,0,0,0,0,1,1,0,0,0,0,0,0},
  {0,0,0,0,0,0,1,1,0,0,0,0,0,0},
  {0,1,0,0,0,0,1,1,0,0,0,0,0,0},
  {0,0,1,1,1,1,1,0,0,0,0,0,0,0},
  {0,0,0,0,0,0,0,0,0,0,0,0,0,0},
  {0,0,0,0,0,0,0,0,0,0,0,0,0,0},
  {0,0,0,0,0,0,0,0,0,0,0,0,0,0},
  {0,0,0,0,0,0,0,0,0,0,0,0,0,0},
  {0,0,0,0,0,0,0,0,0,0,0,0,0,0},
  {0,0,0,0,0,0,0,0,0,0,0,0,0,0},
  {0,0,0,0,0,0,0,0,0,0,0,0,0,0},
  {0,0,0,0,0,0,0,0,0,0,0,0,0,0},
}}

@@

\begin{figure}[htbp]
  \centering
  \caption{Consolas BDF of '\texttt{a}' character with line height}
  \label{fig:letter_a}

@@

  \foreach \x in {0, 1, ..., 13}{
    \foreach \y in {0, 1, ...,18}{
      \def\bit{0} \pgfmathsetmacro{\bit}{\pixelmap_a[mod(\y,19)][mod(\x,14)]}
      \node[box, fill=color\bit] at (\x,\y){};
    }
  }
  \end{tikzpicture}
\end{figure}
